%==============================================================================
% ACKNOWLEDGEMENTS AND DECLARATION
%==============================================================================
\chapter*{Acknowledgements and Declaration}
\addcontentsline{toc}{chapter}{Acknowledgements and Declaration}

\section*{Acknowledgements}

[To be completed by the author upon thesis finalization.]

I wish to express my sincere gratitude to [Supervising Professor Name] for their methodological guidance and intellectual support throughout this research project.

I am deeply thankful to the DNSI team for providing access to the SharePoint corpus, for their availability during the user study, and for the trust they placed in this research.

I also thank the faculty and staff of Aivancity School for Technology and Business for the rigorous academic environment that made this work possible.

Finally, [personal acknowledgements to be completed by the author].

\bigskip

\section*{Declaration}

I, Jean Direl NZE, declare that:

\begin{enumerate}
  \item This thesis is entirely my own work, except where explicitly stated otherwise through appropriate references and citations.
  \item I have read and understood the plagiarism policy of Aivancity School for Technology and Business, and I confirm that this work does not violate that policy.
  \item I have obtained necessary permissions for any third-party material included in this work.
  \item Generative AI tools (specifically Claude, developed by Anthropic) were used as an assistive tool during the writing process. All uses are documented in Annex~E in compliance with Aivancity's policy on generative AI use. All analysis, reasoning, interpretation, and critical contributions are my own intellectual work.
  \item The DNSI organizational context, documents, and user study data are treated in accordance with the confidentiality agreement signed between myself, the DNSI, and Aivancity School for Technology and Business.
\end{enumerate}

\bigskip

\noindent Signed: \underline{\hspace{5cm}} \hspace{2cm} Date: \underline{\hspace{3cm}}

\noindent Jean Direl NZE
