%==============================================================================
% APPENDIX E — DECLARATION OF AI TOOL USE
%==============================================================================
\chapter{Declaration of Generative AI Tool Use}
\label{app:ai-declaration}

In accordance with Aivancity School for Technology and Business's policy on the use of generative AI tools in academic work \citep{AivancityGuide2025}, the following declaration documents all uses of generative AI in the production of this thesis.

\section*{Tool Used}

\textbf{Claude} (claude-sonnet-4-6), developed by Anthropic. Accessed via the Claude Code CLI interface.

\section*{Scope of Use}

Generative AI assistance was used for the following purposes:

\begin{enumerate}
  \item \textbf{Structural drafting:} Initial drafts of thesis chapter structures, section headings, and outline frameworks were generated with AI assistance and subsequently revised, reorganized, and substantially expanded by the author.
  \item \textbf{Literature synthesis support:} AI assistance was used to suggest connections between cited works and to draft initial paragraph structures for the literature review. All factual claims, citations, and analytical judgments were verified and authored by the student.
  \item \textbf{LaTeX formatting:} Table structures, equation formatting, and bibliographic entry formatting were assisted by AI tools.
  \item \textbf{Language editing:} Grammar, sentence structure, and clarity improvements were suggested by AI tools and accepted or modified by the author.
\end{enumerate}

\section*{Scope of Non-Use}

The following elements were not generated by AI and represent the student's own intellectual work:

\begin{itemize}
  \item The research question formulation and hypothesis design
  \item All architectural design decisions and their justifications
  \item The benchmark dataset construction and annotation
  \item The experimental results and their interpretation
  \item The critical analysis of limitations and future work
  \item The code implementation of the AntiGravity system
  \item All final editorial decisions regarding content, emphasis, and argumentation
\end{itemize}

\section*{Compliance Statement}

This declaration complies with rules 1 (full transparency), 2 (systematic reference), 3 (fair use), and 4 (ethics) as specified in Chapter III of the Aivancity PFE Guidelines \citep{AivancityGuide2025}. The use of generative AI in this thesis is assistive and supplementary; the thesis primarily reflects the student's own skills, ideas, and efforts.

\bigskip

\noindent Signed: \underline{\hspace{5cm}} \hspace{2cm} Date: \underline{\hspace{3cm}}

\noindent Jean Direl NZE
