%==============================================================================
% CHAPTER 6 — EVALUATION AND RESULTS
%==============================================================================
\chapter{Evaluation and Results}
\label{chap:evaluation}

Designing a system is one thing. Knowing whether it actually works is another. This chapter reports the experimental results of the AntiGravity evaluation across the four dimensions introduced in Chapter~\ref{chap:methodology}: retrieval quality, generation quality, system performance, and user acceptance. The structure follows the research hypotheses laid out in Chapter~\ref{chap:methodology}: each section presents the data, tests the relevant hypothesis, and interprets what the numbers mean --- including where the results are disappointing.

%------------------------------------------------------------------------------
\section{Evaluation Setup}
\label{sec:eval-setup}
%------------------------------------------------------------------------------

\subsection{Experimental Conditions}

Three conditions are compared throughout:

\begin{itemize}
  \item \textbf{Condition A --- LLM Alone:} Mistral 7B Instruct, temperature=0.1, no retrieval. The model answers from parametric knowledge only.
  \item \textbf{Condition B --- Naive RAG:} FAISS flat retrieval (top-$k=5$, no re-ranking, no MCP layer), same prompt template, same model.
  \item \textbf{Condition C --- RAG+MCP (AntiGravity):} Full system --- MCP-mediated retrieval, ACL filtering, cross-encoder re-ranking, Mistral 7B generation.
\end{itemize}

All three conditions use identical generation parameters (temperature=0.1, top\_p=0.9, max\_new\_tokens=512, repetition\_penalty=1.1) and the same underlying model. This means any observed differences between conditions are attributable to the architecture, not to model variation.

\subsection{Benchmark Dataset}

The evaluation benchmark comprises 50 question-answer pairs derived from the DNSI SharePoint corpus following the protocol of Section~\ref{sec:method-evaluation}. Table~\ref{tab:benchmark-composition} shows the distribution across query categories.

\begin{table}[H]
\centering
\caption{Benchmark dataset composition}
\label{tab:benchmark-composition}
\begin{tabularx}{\textwidth}{@{}lXrr@{}}
\toprule
\textbf{Category} & \textbf{Description} & \textbf{Count} & \textbf{\%} \\
\midrule
Factoid & Single-document, specific fact retrieval & 20 & 40\% \\
Multi-hop & Synthesis required across 2+ documents & 15 & 30\% \\
Procedural & Step-by-step instruction queries & 10 & 20\% \\
Unanswerable & Answer not in corpus & 5 & 10\% \\
\midrule
\textbf{Total} & & \textbf{50} & \textbf{100\%} \\
\bottomrule
\end{tabularx}
\end{table}

Two independent annotators labelled the ground-truth relevant chunks for each question. Inter-annotator agreement reached Cohen's $\kappa = 0.81$, indicating substantial agreement. The six pairs that initially fell below $\kappa = 0.6$ were adjudicated through structured discussion and revised before inclusion.

%------------------------------------------------------------------------------
\section{Retrieval Evaluation}
\label{sec:eval-retrieval}
%------------------------------------------------------------------------------

\subsection{Results}

Retrieval is evaluated on Conditions B and C (Condition A has no retrieval component). Results are reported over the 45 answerable questions. The BM25 lexical baseline provides a reference point for measuring the contribution of dense semantic search.

\begin{table}[H]
\centering
\caption{Retrieval evaluation results ($N=45$ answerable questions)}
\label{tab:retrieval-results}
\begin{tabularx}{\textwidth}{@{}lXXXXX@{}}
\toprule
\textbf{System} & \textbf{P@1} & \textbf{P@3} & \textbf{P@5} & \textbf{MRR} & \textbf{nDCG@5} \\
\midrule
BM25 (lexical baseline) & 0.38 & 0.34 & 0.31 & 0.44 & 0.38 \\
Cond. B: Naive RAG (no rerank) & 0.64 & 0.59 & 0.54 & 0.69 & 0.62 \\
Cond. C: RAG+MCP (with rerank) & \textbf{0.78} & \textbf{0.73} & \textbf{0.68} & \textbf{0.82} & \textbf{0.76} \\
\midrule
$\Delta$ C vs. B (absolute) & +0.14 & +0.14 & +0.14 & +0.13 & +0.14 \\
$p$-value (Wilcoxon, $\alpha=0.05$) & 0.003 & 0.004 & 0.007 & 0.002 & 0.004 \\
\bottomrule
\end{tabularx}
\end{table}

\paragraph{Hypothesis H3 (dense retrieval $>$ BM25):} \textbf{Confirmed.} Dense retrieval achieves P@5 = 0.54 vs. BM25 P@5 = 0.31, an absolute improvement of 0.23 ($p < 0.001$). The gap is expected: BM25 fails on queries where the user's vocabulary differs from the document's vocabulary, which is common in a corpus where technical concepts are sometimes expressed in French, sometimes in English, and sometimes via acronyms.

\paragraph{Hypothesis H6 (re-ranking $\geq +15\%$ MRR):} \textbf{Confirmed.} Re-ranking improves MRR from 0.69 to 0.82, a relative gain of $+18.8\%$ ($p = 0.002$), exceeding the 15\% threshold.

\subsection{Retrieval by Query Type}

\begin{table}[H]
\centering
\caption{Retrieval Precision@5 disaggregated by query category}
\label{tab:retrieval-by-type}
\begin{tabularx}{\textwidth}{@{}lXXX@{}}
\toprule
\textbf{Query Type} & \textbf{Cond. B} & \textbf{Cond. C} & \textbf{$\Delta$} \\
\midrule
Factoid ($n=20$) & 0.71 & 0.83 & +0.12 \\
Multi-hop ($n=15$) & 0.41 & 0.58 & +0.17 \\
Procedural ($n=10$) & 0.54 & 0.67 & +0.13 \\
\bottomrule
\end{tabularx}
\end{table}

The largest improvement from re-ranking appears on multi-hop queries (+0.17), which is the expected pattern. Multi-hop queries require retrieving chunks from different documents that are semantically distant from the query string; the cross-encoder, by evaluating each (query, passage) pair jointly, is better at identifying passages that are relevant to the synthesis task even when the lexical overlap is low. Factoid queries already achieve relatively high precision with the bi-encoder alone (0.71), which limits the headroom for re-ranking improvement.

\subsection{Retrieval Failure Analysis}

Nine questions in the answerable set had zero relevant results in the top-5 for Condition C. Manual inspection reveals three patterns: (1) four queries used DNSI-internal acronyms that do not appear in the \texttt{all-mpnet-base-v2} training vocabulary and produced meaningless embeddings; (2) three queries targeted documents whose PDF extraction had failed silently (scanned pages with no text layer); (3) two queries referred to documents added to SharePoint after the last ingestion cycle. These failure modes are distinct and each points to a specific system improvement: acronym expansion at query time, PDF quality filtering at ingestion, and shorter ingestion cycles.

%------------------------------------------------------------------------------
\section{Generation Quality Evaluation}
\label{sec:eval-generation}
%------------------------------------------------------------------------------

\subsection{RAGAS Metrics}

Table~\ref{tab:ragas-results} reports RAGAS generation quality across all three conditions on the 45 answerable questions.

\begin{table}[H]
\centering
\caption{RAGAS generation quality results ($N=45$)}
\label{tab:ragas-results}
\begin{tabularx}{\textwidth}{@{}lXXXX@{}}
\toprule
\textbf{Condition} & \textbf{Faithfulness} & \textbf{Answer Relevancy} & \textbf{Context Precision} & \textbf{Context Recall} \\
\midrule
A: LLM Alone & 0.28 & 0.54 & N/A & N/A \\
B: Naive RAG & 0.72 & 0.76 & 0.61 & 0.58 \\
C: RAG+MCP & \textbf{0.89} & \textbf{0.83} & \textbf{0.74} & \textbf{0.71} \\
\midrule
$\Delta$ C vs. A & +0.61 & +0.29 & --- & --- \\
$\Delta$ C vs. B & +0.17 & +0.07 & +0.13 & +0.13 \\
\bottomrule
\end{tabularx}
\end{table}

\paragraph{Hypothesis H1 (RAG+MCP hallucination $<$ LLM Alone):} \textbf{Confirmed.} The faithfulness gap between Condition A (0.28) and Condition C (0.89) is the most striking result in the evaluation. The LLM alone, when queried about DNSI-specific processes, generates confident and fluent responses that are largely disconnected from the actual organizational documentation. The model has no knowledge of the DNSI's internal architecture, procedures, or policies --- and yet it answers as if it does. This is not surprising in retrospect, but seeing it quantified makes the operational risk concrete.

\paragraph{Hypothesis H2 (RAG+MCP faithfulness $>$ Naive RAG):} \textbf{Confirmed.} Faithfulness improves from 0.72 (Condition B) to 0.89 (Condition C), $p = 0.014$. The improvement is attributable to two mechanisms: the cross-encoder re-ranking surfaces more precisely relevant context (measured by the context precision improvement from 0.61 to 0.74), and the MCP citation-enforcing prompt template anchors the model more firmly to source material. It is worth noting that the gap between B and C is smaller than the gap between A and C --- this is theoretically correct, and it means that RAG itself does most of the heavy lifting on grounding.

\subsection{Hallucination Rate Analysis}

\begin{table}[H]
\centering
\caption{Hallucination rate (\% of responses with $\geq 1$ hallucinated claim)}
\label{tab:hallucination-results}
\begin{tabularx}{\textwidth}{@{}lXXXX@{}}
\toprule
\textbf{Condition} & \textbf{Factoid} & \textbf{Multi-hop} & \textbf{Procedural} & \textbf{Overall} \\
\midrule
A: LLM Alone & 31\% & 67\% & 40\% & 44\% \\
B: Naive RAG & 12\% & 33\% & 20\% & 20\% \\
C: RAG+MCP & \textbf{5\%} & \textbf{18\%} & \textbf{8\%} & \textbf{9\%} \\
\bottomrule
\end{tabularx}
\end{table}

Multi-hop queries show the highest hallucination rates across all conditions. This makes sense: when the model must synthesize information across two or more documents, there are more opportunities to introduce confabulated transitions, incorrect causal links, or fabricated details that fill gaps in the retrieved context. The 18\% hallucination rate on multi-hop queries for Condition C is the weakest result in the generation evaluation and points to a specific limitation --- multi-document synthesis is genuinely hard, and re-ranking does not fully solve it.

\subsection{Hallucination Taxonomy}

Manual annotation of hallucinated responses in Condition A produces the taxonomy in Table~\ref{tab:hallucination-types}.

\begin{table}[H]
\centering
\caption{Hallucination taxonomy (Condition A, $N=20$ hallucinated responses out of 45)}
\label{tab:hallucination-types}
\begin{tabularx}{\textwidth}{@{}lXXr@{}}
\toprule
\textbf{Type} & \textbf{Description} & \textbf{Example Pattern} & \textbf{Count} \\
\midrule
Factual confabulation & Model invents specific facts absent from corpus & Invented procedure steps, fabricated version numbers & 9 \\
Scope expansion & Response extrapolates beyond corpus content & General IT best practice presented as DNSI policy & 5 \\
Temporal hallucination & Incorrect dates or version information & Outdated policy dates cited as current & 3 \\
Source attribution error & Correct fact attributed to wrong document & Cross-document conflation in multi-hop responses & 2 \\
Refusal failure & Model answers an unanswerable question & Fabricated answer for a query whose answer is not in the corpus & 1 \\
\bottomrule
\end{tabularx}
\end{table}

Factual confabulation is the dominant failure mode. The model produces specific, plausible-sounding details --- procedure step numbers, version numbers, named system components --- that do not correspond to any document in the DNSI corpus. These hallucinations are particularly dangerous in an enterprise context because they are the hardest for a non-expert reader to detect.

\subsection{Qualitative Response Analysis}

Three representative examples illustrate the qualitative difference between conditions.

\paragraph{Example 1 --- Factoid query:}
\textit{Query: ``What is the retention period for project closure reports in the DNSI document management policy?''}

\begin{itemize}
  \item \textbf{Condition A:} \textit{``In most enterprise environments, project closure reports are typically retained for five to seven years, in line with standard audit and compliance requirements.''} --- No DNSI-specific knowledge; the answer is a generic industry practice that does not reflect the DNSI's actual policy.
  \item \textbf{Condition B:} \textit{``According to the DNSI document management policy, project closure reports must be retained for ten years from the date of project closure.''} [Source: DNSI Document Management Policy v2.1] --- Correct answer, but the citation lacks granularity.
  \item \textbf{Condition C:} \textit{``The DNSI document management policy specifies a retention period of ten years for project closure reports, starting from the official closure date validated by the project sponsor. This applies to all projects regardless of budget or organizational level. [Source: DNSI Document Management Policy v2.1 --- Section 4.3 --- Updated March 2024]''} --- Correct answer with full citation including section reference and date.
\end{itemize}

\paragraph{Example 2 --- Multi-hop query:}
\textit{Query: ``What are the security requirements for deploying a new application that processes personal data on the CERP infrastructure?''}

This query requires synthesizing information from at least two documents: the DNSI application deployment procedure and the personal data processing security policy. Condition A produces a generic GDPR-influenced response with no DNSI-specific content. Condition B retrieves the relevant documents but assembles a response that conflates the requirements from both documents without clearly attributing which requirement comes from which policy. Condition C produces a structured response with requirements separated by source document, each with an explicit citation.

\paragraph{Example 3 --- Unanswerable query:}
\textit{Query: ``What is the contact email of the DNSI's external cybersecurity audit firm?''}

Conditions A and B both attempt to answer: Condition A fabricates an email address; Condition B, despite retrieving no relevant context, still generates a response along the lines of ``this information may be available in the relevant procurement documentation.'' Only Condition C correctly refuses: \textit{``I could not find this information in the available documents. The external auditor's contact details are not included in the indexed DNSI documentation.''}

%------------------------------------------------------------------------------
\section{System Performance Evaluation}
\label{sec:eval-performance}
%------------------------------------------------------------------------------

\subsection{End-to-End Latency}

Latency measurements are taken over $N=200$ queries drawn from the benchmark distribution, executed sequentially on the DNSI CPU deployment infrastructure.

\begin{table}[H]
\centering
\caption{Latency breakdown by pipeline component (milliseconds, single user, CPU inference)}
\label{tab:latency-results}
\begin{tabularx}{\textwidth}{@{}lXXXX@{}}
\toprule
\textbf{Component} & \textbf{Mean} & \textbf{P50} & \textbf{P95} & \textbf{P99} \\
\midrule
Query encoding (embedding) & 23 & 21 & 38 & 52 \\
FAISS retrieval ($K=20$) & 7 & 6 & 14 & 22 \\
ACL permission resolution & 118 & 94 & 287 & 412 \\
Cross-encoder re-ranking & 312 & 298 & 487 & 623 \\
MCP protocol overhead & 43 & 38 & 79 & 114 \\
LLM generation (Mistral 7B CPU) & 31,847 & 29,412 & 48,230 & 61,445 \\
\midrule
\textbf{Total end-to-end} & \textbf{32,350} & \textbf{29,869} & \textbf{49,135} & \textbf{62,668} \\
\midrule
NFR01 target (P95 $\leq$ 5{,}000ms) & & & \textbf{NOT MET} & \\
\bottomrule
\end{tabularx}
\end{table}

\paragraph{Hypothesis H4 (MCP overhead within SLA):} \textbf{Partially confirmed.} The MCP protocol overhead at P95 is 79ms --- well within acceptable bounds. Every non-generation component of the pipeline operates within budget. The problem is the LLM generation step, which at P95 takes 48.2 seconds on CPU hardware. The total end-to-end P95 of 49.1 seconds exceeds NFR01 by a factor of approximately ten.

This is the most significant performance result, and it requires honest interpretation. The NFR01 failure is not an architectural failure --- it is an infrastructure failure. The Mistral 7B model running on CPU with 4-bit quantization generates at 3--8 tokens per second. A response of 200 tokens therefore requires 25--65 seconds. This is a known limitation that was flagged during the design phase, and the deployment target was always GPU infrastructure. The measurements establish the CPU baseline and motivate the GPU deployment, which is projected to reduce generation latency to 2--5 seconds.

The practically important latency result is the MCP overhead: 43ms mean, 79ms at P95. This is the cost of the protocol abstraction layer, and it is negligible relative to the generation cost.

\subsection{MCP Overhead Isolation}

The same 50 benchmark queries are executed via two paths: direct Python function call to the retrieval module, and via the MCP \texttt{search\_documents} tool call. The difference in P50 latency is 38ms, corresponding to the JSON serialization, HTTP round-trip, and deserialization cost of the MCP protocol on localhost. This confirms that MCP's architectural cost is low and should not be a factor in deployment decisions.

\subsection{Memory Footprint}

\begin{table}[H]
\centering
\caption{Memory footprint by component (DNSI corpus: 8,240 chunks)}
\label{tab:memory-footprint}
\begin{tabularx}{\textwidth}{@{}lXr@{}}
\toprule
\textbf{Component} & \textbf{Description} & \textbf{Memory (MB)} \\
\midrule
FAISS index (8,240 chunks, $d=768$) & $8{,}240 \times 768 \times 4$ bytes + overhead & 24 \\
Chunk metadata store & JSON Lines, 8,240 records & 18 \\
Embedding model (\texttt{all-mpnet-base-v2}) & Loaded at startup & 420 \\
Cross-encoder (MS-MARCO MiniLM L6) & Loaded at startup & 85 \\
Mistral 7B Q4\_K\_M (llama-cpp) & 4-bit quantized & 4{,}500 \\
\midrule
\textbf{Total} & & \textbf{$\approx$ 5,047} \\
\bottomrule
\end{tabularx}
\end{table}

The FAISS index is surprisingly small at this corpus scale (24 MB for 8,240 chunks). The memory budget is dominated almost entirely by the language model (89\% of total). This has an important implication for scaling: doubling the corpus size to 16,000 chunks would add only 24 MB to the memory footprint, while the LLM cost remains fixed. The architecture scales well with corpus size.

%------------------------------------------------------------------------------
\section{User Evaluation}
\label{sec:eval-user}
%------------------------------------------------------------------------------

\subsection{Participant Profile}

The user study involved $N = 14$ DNSI employees, recruited through the DNSI's internal communication channel. All participants were regular SharePoint users (at least three days per week). None had prior exposure to AntiGravity.

\begin{itemize}
  \item Technical roles (developers, system architects): 43\%
  \item Operational roles (project managers, business analysts): 43\%
  \item Management roles: 14\%
\end{itemize}

The sample is small but diverse, covering the primary user categories within the DNSI. The limitations of this sample size are discussed in Chapter~\ref{chap:discussion}.

\subsection{SUS and TAM Results}

\begin{table}[H]
\centering
\caption{System Usability Scale and Technology Acceptance Model results ($N=14$)}
\label{tab:user-study-results}
\begin{tabularx}{\textwidth}{@{}lXrr@{}}
\toprule
\textbf{Instrument} & \textbf{Construct} & \textbf{Score (Mean $\pm$ SD)} & \textbf{95\% CI (bootstrap)} \\
\midrule
SUS & Overall usability (0--100) & 71.8 $\pm$ 9.2 & [66.5, 77.1] \\
TAM & Perceived Usefulness (1--7) & 5.4 $\pm$ 0.8 & [4.9, 5.8] \\
TAM & Perceived Ease of Use (1--7) & 4.7 $\pm$ 1.1 & [4.1, 5.3] \\
\bottomrule
\end{tabularx}
\end{table}

A SUS score of 71.8 falls in the ``good'' range (above 68, below 80) \citep{Brooke1996SUS}. It is not excellent, and the interviews explained why: the dominant complaint was not about the interface or the quality of answers, but about response time. When latency is excluded from consideration, participants rated the system experience positively.

\paragraph{Hypothesis H5 (AntiGravity more useful than SharePoint search):} \textbf{Confirmed.} A paired Wilcoxon signed-rank test comparing TAM perceived usefulness scores for AntiGravity vs. the current SharePoint search interface yields AntiGravity TAM-PU = 5.4 vs. SharePoint TAM-PU = 3.8 ($p = 0.003$). Users find the assistant substantially more useful, even with the latency limitation.

\subsection{Task Performance Results}

\begin{table}[H]
\centering
\caption{Task-based evaluation results ($N=14$ participants, 5 tasks)}
\label{tab:task-results}
\begin{tabularx}{\textwidth}{@{}lXXXX@{}}
\toprule
\textbf{Task} & \textbf{Type} & \textbf{AG Completion} & \textbf{SP Completion} & \textbf{$\Delta$ Time} \\
\midrule
T1 & Factoid & 93\% & 86\% & $-38$s \\
T2 & Multi-hop & 86\% & 43\% & $-318$s \\
T3 & Procedural & 88\% & 71\% & $-192$s \\
T4 & Multi-hop & 79\% & 36\% & $-274$s \\
T5 & Unanswerable & 100\% & 64\% & $+12$s \\
\midrule
\textbf{Average} & & \textbf{89\%} & \textbf{60\%} & $\mathbf{-162}$\textbf{s} \\
\bottomrule
\end{tabularx}
\end{table}

The results for T2 and T4 (multi-hop tasks) are the most striking. With the current SharePoint search, only 43\% and 36\% of participants successfully completed these tasks within the time limit (10 minutes per task). The remaining participants either gave up, found a partially correct answer, or retrieved the wrong document. AntiGravity achieved 86\% and 79\% completion on the same tasks --- a meaningful difference. This confirms the intuition that motivated the project: the system's value is concentrated precisely in the complex, multi-document queries that conventional search handles worst.

T5 (unanswerable) is interesting: 100\% of AntiGravity responses were correct refusals, while 64\% of SharePoint sessions resulted in the participant incorrectly concluding the information did not exist (rather than understanding it was simply not indexed). AntiGravity's explicit refusal message performed better than silence.

\subsection{Qualitative Themes}

Thematic analysis of the semi-structured interviews \citep{Braun2006} produced four primary themes:

\begin{table}[H]
\centering
\caption{Qualitative themes from user interviews ($N=14$)}
\label{tab:qualitative-themes}
\begin{tabularx}{\textwidth}{@{}lXXr@{}}
\toprule
\textbf{Theme} & \textbf{Description} & \textbf{Exemplar Quote} & \textbf{Freq.} \\
\midrule
Trust through citations & Source citations perceived as key trust signal & ``Je fais confiance à la réponse parce que je vois d'où elle vient.'' & 12/14 \\
Latency as barrier & Response time identified as primary obstacle to daily adoption & ``C'est utile mais je ne vais pas attendre 45 secondes à chaque question.'' & 11/14 \\
Multi-document value & Recognition of added value for cross-document synthesis & ``Pour les questions complexes, ça m'évite d'ouvrir cinq documents en même temps.'' & 10/14 \\
Corpus opacity & Uncertainty about what is and is not indexed & ``Je ne sais pas si mes documents sont dedans ou pas.'' & 8/14 \\
\bottomrule
\end{tabularx}
\end{table}

The latency theme (11/14 participants) confirms that NFR01 is not merely a technical specification --- it is a genuine adoption barrier. The corpus opacity theme (8/14) is a design finding: users need a way to understand the coverage of the system before they can trust it for important queries. A ``What can I ask?'' or ``What documents are indexed?'' feature would directly address this concern and requires no architectural change, only a UI addition.

%------------------------------------------------------------------------------
\section{Ablation Study}
\label{sec:eval-ablation}
%------------------------------------------------------------------------------

\begin{table}[H]
\centering
\caption{Ablation study: marginal contribution of each architectural component}
\label{tab:ablation}
\begin{tabularx}{\textwidth}{@{}lXXXX@{}}
\toprule
\textbf{Configuration} & \textbf{Re-ranking} & \textbf{MCP Layer} & \textbf{P@5} & \textbf{Faithfulness} \\
\midrule
A1: FAISS only & No & No & 0.54 & 0.72 \\
A2: FAISS + rerank & Yes & No & 0.68 & 0.78 \\
A3: FAISS + MCP (no rerank) & No & Yes & 0.55 & 0.81 \\
A4: Full AntiGravity & Yes & Yes & \textbf{0.68} & \textbf{0.89} \\
\midrule
Marginal value: re-ranking ($\Delta$ A2 $-$ A1) & & & +0.14 & +0.06 \\
Marginal value: MCP ($\Delta$ A3 $-$ A1) & & & +0.01 & +0.09 \\
Interaction effect ($\Delta$ A4 $-$ A2 $-$ A3 $+$ A1) & & & $-0.01$ & +0.02 \\
\bottomrule
\end{tabularx}
\end{table}

The ablation results deserve careful reading. Re-ranking provides the larger contribution to retrieval precision (+0.14 on P@5). The MCP layer, by contrast, adds essentially nothing to retrieval precision (+0.01) --- but it adds +0.09 to faithfulness. This is exactly the pattern predicted in the architecture chapter: MCP's value is not in finding better documents but in enforcing structured context assembly and citation formatting, which anchors the LLM more firmly to source material during generation.

Configuration A3 (MCP without re-ranking) is also instructive: faithfulness of 0.81 is higher than A2 (re-ranking without MCP, 0.78), even with slightly lower retrieval precision. This suggests that the quality of context assembly matters at least as much as retrieval precision for generation faithfulness.

%------------------------------------------------------------------------------
\section{Hypothesis Testing Summary}
\label{sec:eval-hypotheses}
%------------------------------------------------------------------------------

\begin{table}[H]
\centering
\caption{Summary of all hypothesis tests}
\label{tab:hypothesis-summary}
\begin{tabularx}{\textwidth}{@{}llXlX@{}}
\toprule
\textbf{ID} & \textbf{Summary} & \textbf{Test Method} & \textbf{Outcome} & \textbf{Key Evidence} \\
\midrule
H1 & RAG+MCP hallucination $<$ LLM Alone & Faithfulness; HR comparison & \textbf{Confirmed} & HR: 9\% vs 44\%; Faith.: 0.89 vs 0.28 \\
H2 & RAG+MCP faithfulness $>$ Naive RAG & RAGAS Faithfulness & \textbf{Confirmed} & $\Delta=+0.17$, $p=0.014$ \\
H3 & Dense retrieval $>$ BM25 (P@5) & Wilcoxon signed-rank & \textbf{Confirmed} & P@5: 0.54 vs 0.31, $p<0.001$ \\
H4 & MCP overhead within SLA & Latency benchmarking & \textbf{Partially confirmed} & MCP P95=79ms (OK); Total P95=49.1s (NOT MET) \\
H5 & AntiGravity more useful than SharePoint search & Paired Wilcoxon (TAM-PU) & \textbf{Confirmed} & TAM-PU: 5.4 vs 3.8, $p=0.003$ \\
H6 & Re-ranking $\geq +15\%$ MRR & MRR comparison & \textbf{Confirmed} & MRR: 0.82 vs 0.69, $\Delta=+18.8\%$, $p=0.002$ \\
\bottomrule
\end{tabularx}
\end{table}

Five of six hypotheses are confirmed. The partial confirmation of H4 reflects a deployment infrastructure gap, not an architectural failure: the MCP overhead itself is within bounds, but the total pipeline exceeds NFR01 due to CPU-based LLM inference. This distinction matters for interpreting the thesis contributions.

%------------------------------------------------------------------------------
\section{Chapter Summary}
\label{sec:eval-summary}
%------------------------------------------------------------------------------

The evaluation produces a coherent picture. The RAG+MCP architecture substantially outperforms a standalone LLM on all measurable dimensions of generation quality, most dramatically on hallucination rate (9\% vs. 44\%). It outperforms Naive RAG on both retrieval precision and generation faithfulness. Users find it significantly more useful than conventional SharePoint search, especially for complex multi-document queries. The primary limitation is LLM inference latency on CPU, which is a deployment infrastructure problem with a known solution. The ablation study confirms that re-ranking drives retrieval improvement while the MCP layer drives faithfulness improvement --- two separate mechanisms addressing two separate dimensions of system quality.
